---
title: "Analyze U.S. Census Data using Tidycensus"
author: "Yifei Liu"
date: '2018-11-01'
description: Desc
hacker_news_id: ''
lobsters_id: ''
meta_img: /images/image.jpg
slug: analyze-u-s-census-data-using-tidycensus
tags:
- cesus
- tidycensus
categories: Data Visualization
---

Just fininsh [Tidycensus](https://walkerke.github.io/tidycensus/) class on Datacamp which is called [Analyzing US Census Data in R](https://www.datacamp.com/courses/analyzing-us-census-data-in-r) so I think it will will a great opportunity for me to practice my data visualization skills.

## Show Pennsylvania Median Income 

You need to obtain a API code from [U.S. Census](https://api.census.gov/data/key_signup.html) and set it up in .Rprofile. 

```{r, warning= F, message = F}
library(tidyverse)
library(tidycensus)
library(ggplot2)
library(purrr)
library(dplyr)
library(shiny)
library(viridis)
library(tmap)
library(sf)
library(plotly)
# devtools::install_github('thomasp85/gganimate')
library(gganimate)
# census_api_key(key = "Your API Keys")
library(ggthemes)
library(mapview)
```
Next we need to find median income data, you can use load_variables function to find the variables name that you want to find, you can also use [Census Reporter](https://censusreporter.org) to browser different categories of data. 

```{r, warning= F, message = F}
pa_medianincome <- map_df(2012:2016, function(x){
  get_acs(geography = "county",
          variables = c(Median_income = "B19013_001"),
          state = "PA",
          survey = "acs1",
          year = x) %>%
    mutate(year = x)
})

pa_medianincome_sim <- pa_medianincome %>%
  mutate(NAME = str_replace(NAME, " County, Pennsylvania", ""))

g <- ggplot(pa_medianincome_sim, aes(x = estimate, y = reorder(NAME, estimate))) + 
  geom_errorbarh(aes(xmin = estimate - moe,
                     xmax = estimate + moe,
                     frame = year, ids = NAME))
  geom_point(size = 3, color = "darkgreen", aes(frame = year, ids = NAME))
ggplotly(g)
```

## lets take a look of U.S. poverty 



```{r,warning= F, message = F}
us_pov <- get_acs(geography = "county",
                  variables = "B17001_002",
                  summary_var = "B17001_001",
                  geometry = T, 
                  shift_geo = T) %>%
  mutate(pov_pct = 100*(estimate/summary_est))
us_pov
```

Here it is, the U.S. poverty rate in each counties in U.S. You can use _map_ function to get more year data if you want. But poverty line is around $12,488 for one person and $15,877 for 2 people according to [U.S. Census Bureau](https://www.census.gov/data/tables/time-series/demo/income-poverty/historical-poverty-thresholds.html) in 2017. I normally don't associated U.S. with the world poverty, but after spend some time read articles from magazine such as Foreign Affairs and read wikipedia. I found out American actually spend a ton of money on [war on poverty](https://en.wikipedia.org/wiki/War_on_Poverty), which program start under president Lyndon B. Johnson in 1964 and cost around **$15 trillion**. And the poverty rate still hover around 10% ~ 15%.  


```{r}
ggplot(us_pov) +
  geom_sf(aes(fill = pov_pct), color = NA) +
  scale_fill_viridis(direction = -1) +
  theme_minimal() +
  labs(title = "Percentage of Population Live Below Poverty Line",
       subtitle = "Alaska and Hawaii are being shifted and are not to scale",
       caption = "Source: 2012-2016 5-year ACS",
       fill = "% in Poverty")
```


## Philly House Occupied rate

Let's take a loot at house occupied rate in Philadelphia. 

```{r, warning= F, message = F}
philly_house_pct <- get_acs(geography = "tract",
                       variables = c(Occupied = "B25002_002"),
                       summary_var = "B25002_001",
                       state = "PA",
                       county = "Philadelphia",
                       geometry = T,
                       year = 2016,
                       cb = F) %>%
  mutate(Occupied_pct = 100 * estimate / summary_est)

st_erase <- function(x, y) {
  st_difference(x, st_union(st_combine(y)))
}

philly_water <- area_water("PA", "Philadelphia", year = 2016, class = "sf")
philly_house_pct_noriver <- st_erase(philly_house_pct, philly_water)
mapview(philly_house_pct_noriver, zcol = "Occupied_pct", legend = T)
```


## Metropolitan area rent to income ratio
Largest U.S. metropolitan are is [New York-Newark-Jersey City, NY-NJ-PA](https://en.wikipedia.org/wiki/New_York_metropolitan_area) has population over 20
millions, we can take a look at its rent to income ratio to see what percentage of income people need to spend on its rent. 

```{r, warning= F, message = F}
household_income <- get_acs(geography = "tract",
                        variables = c(income = "B19013_001",
                                      rent = "B25064_001"),
                        state = c("NY", "NJ", "PA"),
                        output = "wide",
                        geometry = T) 

household_pct <- household_income %>%
  mutate(rent = 12 * rentE,
         rent_income = rent/incomeE * 100) %>%
  select(GEOID, NAME, rent_income)
  

metros <- core_based_statistical_areas(cb = T) %>%
  dplyr::filter(GEOID %in% c("10900", "35620", "15380")) %>%
  select(metro_name = NAME)

metros_rent <- st_join(household_pct, metros, join = st_within,
                                 left = FALSE)
head(metros_rent)
```

Now we can explore metropolitan area. We can take a look at kernal density plot of median gross income rent. 

```{r, warning= F, message = F}
ggplot(metros_rent, aes(x = rent_income)) +
  geom_density(position = "identity", alpha = 0.25, aes(fill = metro_name)) +
  labs(x = "rent to income", y = "Density", legend = "Metropolitan",
       fill = "Metro") +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 6))

```

We can also feature geometry as well, we can use _tmap_ to make faceted map. 
```{r, warning= F, message = F}
tm_shape(metros_rent, projection = 26910) +
  tm_polygons("rent_income",title = "", showNA = TRUE,
              border.col = "gray50", border.alpha = .3,
              title = expression("Rent to income")) +
  tm_facets(by = "metro_name", free.coords = T) +
  tm_scale_bar(position = c("left", "bottom")) 
```


## Conclusion

This is my first post, the reason I write this blog is after listen to what David Robinson said in Data Stories, the best way to learn data science is to start a blog, which I strongly agree. I hope this post will be the first step for me to be a data scientist. 


## Reference
- [Spatial data in tidycensus](https://walkerke.github.io/tidycensus/articles/spatial-data.html)
- [Compare US metropolitan area characteristics in R with tidycensus and tigris](https://walkerke.github.io/2017/06/comparing-metros/)
- [Using Tidycensus And Leaflet To Map Census Data](https://juliasilge.com/blog/using-tidycensus/)









