---
title: Leading Cause of Death in the United States
author: Yifei Liu
date: '2018-11-28'
slug: leading-cause-of-death-in-the-united-states
categories:
  - Data Visualization
tags:
  - data
  - R
  - cesus
description: Desc
hacker_news_id: ''
lobsters_id: ''
meta_img: /images/image.jpg
---

In this post I want to review the trend of leading causes of deaths in United States. Specially how fast the age adjusted rate decase in each state. The data come from [CDC](https://data.cdc.gov/NCHS/NCHS-Leading-Causes-of-Death-United-States/bi63-dtpu). The main variable I try to analysis in this post is [age-adjusted death rates](https://health.mo.gov/data/mica/CDP_MICA/AARate.html). 

### Pacakge and Data


```{r, warning=F, message=F}
library(tidyverse)
library(jsonlite)
library(fuzzyjoin)
library(transformr) # devtools::install_github("thomasp85/transformr")
library(sf)
library(gganimate)
library(geofacet)
library(ggrepel)

theme_set(theme_minimal())

death_data <- read_csv("/Users/yifeiliu/Documents/R/data/EDA/leading_death.csv")

death_data <- death_data %>%
  dplyr::rename(year = Year, 
                all_cause = `113 Cause Name`, 
                cause = `Cause Name`, 
                state = State, 
                deaths = Deaths, 
                aadr = `Age-adjusted Death Rate`)


death_data_processed <- death_data %>%
  select(- all_cause) %>%
  dplyr::filter(!cause == "All causes",
                !str_detect(state, "United"))


```

Since we have the state information, we can use *tidycensus* packge to obtain spatial file and merge those togher. 

```{r, warning=F, message = F}
pop_data_2010 <- get_decennial(geography = "state",
                               variables = "P001001",
                               geometry = T,
                               year = 2010,
                               shift_geo = T)

states <- pop_data_2010 %>%
  select(NAME, geometry) %>%
  setNames(c("state", "geometry"))
```


### Explore the dataset


We can take a look at the leading cuase of people death during those years

```{r,warning=F}

death_data_processed %>%
  group_by(year, cause) %>%
  dplyr::summarise(avg_death = mean(aadr)) %>%
  ggplot(aes(year, avg_death, color = cause)) +
  geom_line() +
  labs(y = "Age Adjusted Death Rate",
       subtitle = "per 100,000 population",
       x = "",
       title = "Leading Cause of Death In U.S. 1999-2016")


death_data %>%
  mutate(year = as.integer(year)) %>%
  dplyr::filter(!cause == "All causes") %>%
  group_by(year, cause)  %>%
  dplyr::summarise(ave_death = mean(aadr)) %>%
  ggplot(aes(ave_death, fct_reorder(cause, ave_death))) +
  geom_point() +
  labs(y = "Leading Death",
       x = "Death Number",
       title = "Leading Cause of Death in U.S. in year: {frame_time}",
       subtitle = "Deaths are per 100,000 population",
       caption = "Data from CDC") +
  transition_time(year) +
  ease_aes('linear')


```



### Visualize change in age adjusted death rate

Aggregated across state and year, see which state death rate increase or decrease


```{r}

death_data_spread <- death_data_processed %>%
  dplyr::mutate(year = paste0("Y", year)) %>%
  group_by(year, state) %>%
  dplyr::summarize(avg_death = mean(aadr)) %>%
  dplyr::mutate(i = row_number()) %>%
  spread(year, avg_death) %>%
  select(-i) %>%
  mutate(current = Y2016,
         change = Y2016 - Y1999)


death_data_spread  %>%
  select(state, current, change) %>%
  ggplot(aes(current, change)) +
  geom_point() +
  geom_text(aes(label = state), vjust = 1, hjust = 1, check_overlap = T)


```


We can visualize the chagne see which place in United State have the highest AADR drop. 


```{r, warning=F, message=F}
death_data_spread  %>%
  select(state, change) %>%
  inner_join(states, by = "state") %>%
  ggplot() +
  geom_sf(aes(fill = change)) +
  coord_sf(crs = 26910) +
  labs(title = "Change of Death Rate",
       subtitle = "1999 : 2016 Death Rate Decrease by",
       caption = "Death rate per 100,000 people") +
  scale_fill_viridis_c(direction = -1) +
  theme_minimal()

death_data_processed %>%
  group_by(year, state) %>%
  dplyr::summarise(avg_death = mean(aadr)) %>%
  mutate(year = as.integer(year)) %>%
  inner_join(states, by = "state") %>%
  ggplot() +
  geom_sf(aes(fill = avg_death)) +
  coord_sf(crs = 26910) +
  labs(title = "Death rate in U.S. year: {frame_time}",
       subtitle = "Data from CDC",
       caption = "Death rate per 100,000 people") +
  scale_fill_viridis_c(direction = -1) +
  transition_time(year)

death_data_processed %>%
  group_by(year, state) %>%
  dplyr::summarise(avg_death = mean(aadr)) %>%
  mutate(year = as.integer(year)) %>%
  ggplot(aes(year, avg_death)) +
  geom_line() +
  scale_x_continuous(labels = function(x) paste0("'", substr(x, 3, 4))) +
  facet_geo(~ state, grid = "us_state_grid2", label = "name") +
  labs(x = "",
       y = "",
       title = "Age Adjusted Death Rate in United States",
       caption = "@yifeiliu Source: Center for Disease Control and Prevention")

```


### The End

After visualize the data, we can draw some conclusion.

* the main causes of death are _heart disease_ and _cancer_.
* age adjusted death rate is decline in all states from 1999 to 2016
* some state age adjusted death rate better the soem other state. 





